### 2017-08-01 paper.tex
Selva add

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint       Remove this option only once the paper is in final form.
%  9pt           Set paper in  9-point type (instead of default 10-point)
% 11pt           Set paper in 11-point type (instead of default 10-point).
% numbers        Produce numeric citations with natbib (instead of default author/year).
% authorversion  Prepare an author version, with appropriate copyright-space text.

\usepackage{amsmath}

\newcommand{\cL}{{\cal L}}
\newcommand{\langname}{Trinity}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
\copyrightdoi{nnnnnnn.nnnnnnn}

% For compatibility with auto-generated ACM eRights management
% instructions, the following alternate commands are also supported.
%\CopyrightYear{2016}
%\conferenceinfo{CONF'yy,}{Month d--d, 20yy, City, ST, Country}
%\isbn{978-1-nnnn-nnnn-n/yy/mm}\acmPrice{\$15.00}
%\doi{http://dx.doi.org/10.1145/nnnnnnn.nnnnnnn}

% Uncomment the publication rights used.
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}  % default
%\setcopyright{rightsretained}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

%\title{Title Text\titlenote{with optional title note}}
\title{\langname: A Language for Multi-View Architecture Description and Control}
%\subtitle{Subtitle Text, if any\titlenote{with optional subtitle note}}

%\authorinfo{Name1\thanks{with optional author note}}
%           {Affiliation1}
%           {Email1}
%\authorinfo{Name2 \and Name3\thanks{with optional author note}}
%           {Affiliation2/3}
%           {Email2/3}
\authorinfo{Jonathan Aldrich}
           {Carnegie Mellon University}
           {jonathan.aldrich@cs.cmu.edu}
\authorinfo{Madaline Kirwin}
           {Grinnell College}
           {kirwinma@grinnell.edu}
\authorinfo{Selva Samuel}
           {Carnegie Mellon University}
           {ssamuel@cs.cmu.edu}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

% 2012 ACM Computing Classification System (CSS) concepts
% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010124.10010138.10010143</concept_id>
<concept_desc>Theory of computation~Program analysis</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Theory of computation~Program analysis}
% end generated code

% Legacy 1998 ACM Computing Classification System categories are also
% supported, but not recommended.
%\category{CR-number}{subcategory}{third-level}[fourth-level]
%\category{D.3.0}{Programming Languages}{General}
%\category{F.3.2}{Logics and Meanings of Programs}{Semantics of Programming Languages}[Program analysis]

\keywords
keyword1, keyword2

%\section{Introduction}

%The text of the paper begins here.

\section{Multiple Views of Software Architecture}

\begin{itemize}
\renewcommand\labelitemi{--}
\item Give the SEI definition of software architecture
\item State the usefulness of software architecture in analysis and design activities
\item Introduce ADLs.
\item Introduce views. State their usefulness in dealing with complexity by separating concerns
\item State the lack of support for multiple views in ADLs.
\item State that industry practitioners desire support for multiple views. Cite study by Malavolta et. al.
\end{itemize}
The software architecture of a system is the set of structures needed to reason about the system, which comprise software elements, relations among them, and properties of both [].

%Most software architectures are implicit.

%UML
\cite{Malavolta2013}

\section{Architectural Control}
\begin{itemize}
\renewcommand\labelitemi{--}
\item What is architectural control? --$>$ Architectural constraints must be enforced in the implementation
\item Introduce communication integrity
\item Describe ArchJava's support for enforcing communication integrity in a single JVM and its failure to do so across multiple JVMs
\end{itemize}
\section{\langname}

%\appendix
%\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% The 'abbrvnat' bibliography style is recommended.

\bibliographystyle{abbrvnat}
\bibliography{references.bib}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\end{document}