### 2017-08-12 paper.tex
Everything I have written till now

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[numbers,preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint       Remove this option only once the paper is in final form.
%  9pt           Set paper in  9-point type (instead of default 10-point)
% 11pt           Set paper in 11-point type (instead of default 10-point).
% numbers        Produce numeric citations with natbib (instead of default author/year).
% authorversion  Prepare an author version, with appropriate copyright-space text.

\usepackage{amsmath}
\usepackage{caption}
\usepackage{color}
\usepackage{listings}
\usepackage{url}

\newcommand{\cL}{{\cal L}}
\newcommand{\langname}{Trinity}
\newcommand{\moduleview}{module}
\newcommand{\Moduleview}{Module}

\definecolor{gray}{RGB}{128, 128, 128}

\lstdefinestyle{modulestyle}{
	basicstyle=\scriptsize
}

\lstdefinestyle{cncstyle}{
	aboveskip=0pt,
	belowskip=-10pt,
	basicstyle=\footnotesize\ttfamily,
	captionpos=b,
	columns=flexible,
	float=t,
	frame=single,
	morekeywords={component},
	keywordstyle=\color{blue},
	numbers=left,
	numberstyle=\color{gray}\scriptsize\fontfamily{cmr}\selectfont,
	numbersep=8pt,
	tabsize=2,
	xleftmargin=19pt,
	framexleftmargin=15pt,
	framexrightmargin=-4pt
}

\lstdefinelanguage{cnclang}{
	morekeywords={component, port},
}

\lstdefinestyle{allocationstyle}{
}

\DeclareCaptionFormat{listing}{\textbf{#1.} #3}
\captionsetup[lstlisting]{format=listing}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
\copyrightdoi{nnnnnnn.nnnnnnn}

% For compatibility with auto-generated ACM eRights management
% instructions, the following alternate commands are also supported.
%\CopyrightYear{2016}
%\conferenceinfo{CONF'yy,}{Month d--d, 20yy, City, ST, Country}
%\isbn{978-1-nnnn-nnnn-n/yy/mm}\acmPrice{\$15.00}
%\doi{http://dx.doi.org/10.1145/nnnnnnn.nnnnnnn}

% Uncomment the publication rights used.
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}  % default
%\setcopyright{rightsretained}

%\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

%\title{Title Text\titlenote{with optional title note}}
\title{\langname: A Language for Multi-View Architecture Description and Control}
%\subtitle{Subtitle Text, if any\titlenote{with optional subtitle note}}

%\authorinfo{Name1\thanks{with optional author note}}
%           {Affiliation1}
%           {Email1}
%\authorinfo{Name2 \and Name3\thanks{with optional author note}}
%           {Affiliation2/3}
%           {Email2/3}
%\authorinfo{Selva Samuel}
%           {Carnegie Mellon University}
%           {ssamuel@cs.cmu.edu}
%\authorinfo{Madaline Kirwin}
%           {Grinnell College}
%           {kirwinma@grinnell.edu}
%\authorinfo{Jonathan Aldrich}
%           {Carnegie Mellon University}
%           {jonathan.aldrich@cs.cmu.edu}
\authorinfo{}{}{}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

% 2012 ACM Computing Classification System (CSS) concepts
% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10011006.10011008</concept_id>
%<concept_desc>Software and its engineering~General programming languages</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10003752.10010124.10010138.10010143</concept_id>
%<concept_desc>Theory of computation~Program analysis</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Theory of computation~Program analysis}
% end generated code

% Legacy 1998 ACM Computing Classification System categories are also
% supported, but not recommended.
%\category{CR-number}{subcategory}{third-level}[fourth-level]
%\category{D.3.0}{Programming Languages}{General}
%\category{F.3.2}{Logics and Meanings of Programs}{Semantics of Programming Languages}[Program analysis]

%\keywords
%keyword1, keyword2

%\section{Introduction}

%The text of the paper begins here.

\section{Multiple Views of Software Architecture}

%\begin{itemize}
%\renewcommand\labelitemi{--}
%\item Give the SEI definition of software architecture
%\item State the usefulness of software architecture in analysis and design activities
%\item Introduce ADLs.
%\item Introduce views. State their usefulness in dealing with complexity by separating concerns
%\item State the lack of support for multiple views in ADLs.
%\item State that industry practitioners desire support for multiple views. Cite study by Malavolta et. al.
%\end{itemize}
The software architecture of a system is the set of structures needed to reason about the system, each consisting of elements of the system, relations among them, and properties of both \cite{SEISwArchDef}. Software architecture enables the system designers to analyze the system's ability to meet quality objectives like performance, reliability, etc. Thus, it facilitates the early detection of design errors and leads to improved software quality. It also provides a blueprint for system implementation and evolution.

To facilitate the different uses of software architecture in system development activities, a number of architecture description languages (ADLs) have been developed \cite{ADLs}. ADLs are formal languages that can be used to represent the architecture of a software system in an unambiguous way. Thus, they provide a rigorous basis for the analysis of system designs. ADLs also enable the various stakeholders to understand the system better and thus aid in system implementation, maintenance and evolution.

System designers often organize the architecture description of a software system as a set of views. An architectural view is a representation of those elements of the system that are needed to show how the architecture addresses a concern held by one or more stakeholders. For example, to devise an implementation plan, the system designers might create a (\moduleview) view that shows the module decomposition of the system and the dependencies among the modules (i.e., implementation units). Moreover, this view would not include the system's runtime entities or elements from the system's deployment environment. The separation of concerns achieved by splitting the architectural description into different views helps keep the architecture description cognitively tractable.

The vast majority of existing ADLs, however, lack support for multiple views. This hampers the use of ADLs for the specification of software architecture in practical settings. A survey conducted by \citeauthor{Malavolta2013} \cite{Malavolta2013} found that better support for multiple views is one of the features most desired by industrial practitioners in ADLs.

%This is a limitation. Low expressiveness.

%Software architecture has to address multiple concerns. Using a single model to reason about every concern would make it complex. To deal with this complexity, views are used.

%A view is a representation of one or more structural aspects of an architecture that illustrates how the architecture addresses one or more concerns held by one or more of its stakeholders.
%Most software architectures are implicit.

%UML

\section{Architectural Control}
%\begin{itemize}
%\renewcommand\labelitemi{--}
%\item What is architectural control? --$>$ Architectural constraints must be enforced in the implementation
%\item Introduce communication integrity
%\item Describe ArchJava's support for enforcing communication integrity in a single JVM and its failure to do so across multiple JVMs
%\end{itemize}

Software architects design systems to meet quality attribute requirements like performance, reliability, security, etc. To guarantee that the implemented system exhibits the desired quality attributes, it is necessary to ensure that the implementation adheres to the design principles and constraints prescribed by the architecture. Architectural control is the ability of software architects to enforce architectural constraints on the implementation so that the system meets its design goals.

\citeauthor{LV1995} have identified \textit{communication integrity} as a key aspect of maintaining consistency between the architecture and implementation of a software system \cite{LV1995}. Communication integrity is the property that each component in the implementation may communicate directly only with the components to which it is connected in the architecture.

In prior work, ArchJava used a custom type system to verify communication integrity statically \cite{Aldrich2002}. A limitation of ArchJava is its inability to ensure that an application communicates over the network using only the connections shown in the architecture. Since programs in ArchJava have unrestricted access to system libraries, a component can use the network library directly to communicate in ways not specified in the architecture.

\section{\langname}

We are developing \langname, an ADL that would allow software architects to ...

\noindent
\\
\textbf{Multi-View Architecture Description in \langname.} \citeauthor{Clements2010} \cite{Clements2010} identify three basic types of views: ...

Trinity provides support for these three views.

Module view:

Component-and-connector view:

Deployment view:

\begin{lstlisting}[style=cncstyle, caption=Caption, label=cncview, language=cnclang]
component Test
	port port1
	port port2
\end{lstlisting}

\lstlistingname~\ref{cncview}

\noindent
\\
\textbf{Architectural Control in \langname.}

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

%\acks

%Acknowledgments, if needed.

% The 'abbrvnat' bibliography style is recommended.

\bibliographystyle{abbrvnat}
\bibliography{references.bib}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\end{document}